\documentclass[10pt,oneside,a4paper]{article}
\usepackage{hyperref} 
\usepackage{url} 
\usepackage{cite}



\begin{document}
\title{ Mobile games}
\author{Siarhei Tsikhan\\[2pt]}
\date{\small 8. september 2022}
\maketitle


\section*{History of mobile games}
The early 2000s was a landmark period for the gaming industry - the first games on phones appeared. Mobile games followed the same path as computer games. At first they were primitive arcades, but they also enjoyed great popularity, because it was during this period that phones became more accessible and widespread. For many, it was a pleasant diversion - on the way to work/study, you could not only call, but also play on the phone. One of the most popular games was "Snake" on Nokia phones. Games became more complex and more beautiful in parallel with the development of phones - color displays appeared, power increased, memory increased.\cite{Old}


\section*{Mobile games in the present}
Android games are one of the largest markets with an audience of about 2 billion users. Developers offer both free entertainment apps and paid apps. Among the most famous products are Angry Birds and Pokemon Go.
Android-based applications statistics and features:
\begin{enumerate}
\item Only 3\% of customers donate to mobile games;
\item 80\% of store sales are in game software;
\item In 2010, the volume of mobile game markets was estimated at about 33 billion U.S. dollars;
\item in 2017, the corresponding figure has passed 50 billion (about 43\% of the entire global gaming market);
\item the main consumers are women 35-45 years old.
\end{enumerate}
The most popular genres:
MOBA is a symbiosis of real-time strategy and action. 
Here the user gets to control one character - a champion or a hero.
 His client chooses from a very extensive list initially offered.
The point of the game is to assemble a strong team in order to destroy the enemy's base, as well as to protect their own territory. 
Notable examples of MOBAs are:
\begin{enumerate}
\item Mobile Legends;
\item Arena of Valor;
\item Vainglory.
\end{enumerate}
The computer version of the genre includes Dota 2, Prime World and League of Legends.
 The next popular choice is Battle Royal.
 It is a combination of survival and exploration.
 It brings together a large number of players with minimal equipment.
 They help each other find weapons and collect armor and kill enemies.
The goal is to remain the last survivor. 
The player here needs to compete with competitors by staying as long as possible in safe zones.
 The latter get shallower as time goes on. 
The direction is popular among users 8-16 years old.
Mobile MMORPGs are simplified computer versions.
 They do not require you to spend a huge amount of time learning the mechanics and features of the projects. They consist of:
\begin{enumerate}
\item there are RPG elements;
\item plays a large number of users;
\item The goal is to pump up the character (it is necessary to raise the level of the hero) and his skills;
\item interaction of hundreds of accounts with each other simultaneously within one world is supported;
\item contacting is carried out taking into account the real time mode.
\end{enumerate}
Puzzles are preferred by 50\% of mobile platform users. Logic applications focus on mechanics that are simple to understand, but very difficult to master. They require quick reactions and good intelligence.
Casual games are easy to start. They can be left at any point and then return from where the game was finished. This allows you to spend small amounts of time in the game.\cite{Now}

\section*{The positive impact of playing on a mobile}
Video games allow you to get rid of your thoughts and give your brain a much-needed refresh. Many games require abstract ideas and high-level thinking to break records. These skills are sometimes not even academically relevant.
A good video game session on your cell phone can lift your spirits. The feeling of victory that is realized on the challenging level of a mobile game is incomparable. It usually accompanies games in which users must find their way and destroy enemies. Games like Clash Royale or Crown Towers can bring exciting visuals and fantastic mechanics.
When one is in the virtual world, one rests from the monotony. It is especially useful to keep emotions stable.
Games, especially action genres, require players to make split-second decisions. Games that simulate stressful events, such as battles or combat games, can serve as learning tools for real-life situations. According to the study, these games can help train surgeons and military personnel. Action games force players to make more decisions per unit of time.
Even puzzle or bingo games have significantly improved players' mental activity. Tasks such as keeping track of different bingo cards and focusing on the numbers being called are great for mental stimulation.\cite{Good} 

\section*{The negative impact of playing  on a mobile}
Video games on cell phones have been proven to have a number of side effects. 
Children who play violent video games are aggressive . Children who watch huge amounts of simulated violence may eventually become immune to it. Such people are more likely to self-harm or lack empathy. The American Psychological Association (APA) also found a consistent correlation between violent video games and criminal violence.
Children and adults can become highly addicted to playing video games, In June 2018, the World Health Organization declared gaming addiction a mental disorder. A study by the National Institute of Media and Families in Minneapolis. open that video games can make children addicted. In turn, this leads to high levels of depression and anxiety.
Such children also experience social phobia. Their school performance is at the bottom and they may feel different from their peers. Neuroimaging studies show that heavy use of certain video games can cause radical brain changes.\cite{Bad} 

\bibliography{literatura}
\bibliographystyle{plain} 
\end{document}